---
title: "Dicty analysis"
output:
  html_document: 
    toc: true
    toc_depth: 3
    toc_float: true

  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
#TODO add root folder using knitr root.dir()
```

```{r message = FALSE, warning=FALSE}
library(ggplot2)
library(RVenn)

# load the data and visualization functions
source("global.R")
source("../utils/visualization_functions.R")
```

# summary
Summary statistics Dicty
```{r}
all_data_dicty %>% 
  summarise(Unique_bait = n_distinct(bait))
```

Show the unique baits count
```{r}
all_data_dicty %>% 
  group_by(bait, condition) %>% 
  tidyr::unite(bait_condition, c(bait, condition), sep = '_') %>% 
  summarise(Unique_bait_condition = n_distinct(bait_condition))
```

Show how many instances (lines) of bait there are
```{r}
table(all_data_dicty$bait)
```

Show how many instances (lines) of condition there are
```{r}
table(all_data_dicty$condition)
```

Show total spectral count per bait + condition
```{r}
all_data_dicty %>% 
  group_by(bait, condition) %>% 
  summarise(.groups = 'drop', across(spectral_count, sum))
```

Show normalized spectral count per bait + condition
```{r}
all_data_dicty %>% 
  group_by(bait, condition) %>% 
  summarise(.groups = 'drop', across(z_score_spectral, sum))
# this does not look correct, we should normalize by total spectral count?, can we sum z-scores?
```

Show how many uniq uniprot id's are in the set (dicty has 6 chromosomes and 12500 proteins)
```{r}
n_distinct(all_data_dicty$uniprot)
```

uniq proteins per bait_condition
```{r}
bait_condition <- group_by(all_data_dicty, bait, condition)

bait_condition %>% 
  summarise(n_distinct(uniprot))
```

Unique proteins per condition
```{r}
group_by(all_data_dicty, condition) %>% 
  summarise(n_distinct(uniprot))
```

Unique proteins per bait
```{r}
group_by(all_data_dicty, bait) %>% 
  summarise(n_distinct(uniprot))
```

## absolute spectral count
Barplot showing the total spectral count for each bait
```{r}
p <- ggplot(all_data_dicty, aes(x=bait, y =spectral_count))
p <- p + geom_col() + labs(title = 'Spectral count vs bait')
p
```

Barplot showing the absolute spectral count for each condition
```{r}
p <- ggplot(all_data_dicty, aes(x=condition, y =spectral_count))
p <- p + geom_col() + labs(title = 'Spectral count vs condition')
p
```

Combined spectral count for bait + condition, sum spectral count per group
```{r}
my_data <-all_data_dicty %>% group_by(bait, condition) %>% 
    # .groups = drop is the default, will get warnings if you not specify explicitly 
    summarise(.groups = 'drop', across(spectral_count, sum)) %>% 
    tidyr::unite(bait_condition, c(bait, condition), sep = "_")
  
p <- ggplot(my_data, aes(x = bait_condition, y = spectral_count))
p <- p + geom_col() + labs(title = 'spectral count vs bait + condition')
p + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## average spectra
Barplot showing the average spectral count per protein for each bait
divided the summed group count by the number of observations in that group so avg count per group
```{r}
# get group sizes
group_spectra_normalized <- all_data_dicty %>% 
  group_by(bait) %>% 
  summarise(sum_spectra = sum(spectral_count), group_size = n(), average_spectra = sum(spectral_count/n()))

p <- ggplot(group_spectra_normalized, aes(x=bait, y = average_spectra))
p <- p + geom_col() + labs(title = 'average Spectral count per bait')
p
```

Barplot showing the average spectral count for each condition:
divided the summed group count by the number of observations in that group, 
```{r}

# get group sizes
group_spectra_normalized <- all_data_dicty %>% 
  group_by(condition) %>% 
  summarise(sum_spectra = sum(spectral_count), group_size = n(), average_spectra = sum(spectral_count/n()))


p <- ggplot(group_spectra_normalized, aes(x=condition, y=average_spectra))
p <- p + geom_col() + labs(title = 'average Spectral count vs condition')
p
```

Combined spectral count for bait + condition, average spectral count per group
```{r}
# get group sizes
group_spectra_normalized <- all_data_dicty %>% 
  group_by(bait, condition) %>% 
  summarise(sum_spectra = sum(spectral_count), group_size = n(), average_spectra = sum(spectral_count/n())) %>% 
  tidyr::unite(bait_condition, c(bait, condition), sep = "_")

p <- ggplot(group_spectra_normalized, aes(x = bait_condition, y = average_spectra))
p <- p + geom_col() + labs(title = 'average spectral count vs bait + condition')
p + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Combined spectral count for bait + condition, average spectral count per group
Add overlay with unique proteins 
```{r}
# get group sizes
group_spectra_normalized <- all_data_dicty %>% 
  group_by(bait, condition) %>% 
  summarise(sum_spectra = sum(spectral_count), group_size = n(), normalized_spectra = sum(spectral_count/n()), unique_proteins = n_distinct(uniprot)) %>% 
  tidyr::unite(bait_condition, c(bait, condition), sep = "_")


p <- ggplot(group_spectra_normalized, aes(x = bait_condition, y = normalized_spectra))
p <- p + geom_col() + labs(title = 'average spectral count vs bait + condition') +
geom_text(aes(label=unique_proteins), vjust=0)

p + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## boxplot bait absolute count
Boxplot absolute count vs bait
```{r}
p <- ggplot(all_data_dicty, aes(x=bait, y =log10(spectral_count)))
p <- p + geom_boxplot(outlier.colour = "black") + labs(title = 'Spectral count vs bait')
p
```

## boxplot condition absolute count
Boxplot count vs condition
```{r}
p <- ggplot(all_data_dicty, aes(x=condition, y =log10(spectral_count)))
p <- p + geom_boxplot(outlier.colour = "black") + labs(title = 'Spectral count vs condition')
p
```

## boxplot experiment absolute count
Boxplot combined spectral count for bait + condition
```{r}
my_data <-all_data_dicty %>% 
  group_by(bait, condition) %>% 
    tidyr::unite(bait_condition, c(bait, condition), sep = "_")

my_data$bait_condition <- as.factor(my_data$bait_condition)
p <- ggplot(my_data, aes(x = bait_condition, y = log10(spectral_count)))

p <- p + geom_boxplot(outlier.colour = "black") + labs(title = 'spectral count vs bait + condition')

p + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


<!--
TODO, check why we have the same numbers for some conditions or baits
filter(all_data_dicty, bait == "Gbeta2")
filter(all_data_dicty, bait == "Gbeta1")
Shows the same number of rows retrieved, counts look different for now....
-->

## z-score normalized boxplot count vs bait
```{r}
p <- ggplot(all_data_dicty, aes(x=bait, y =log2(z_score_spectral)))
p <- p + geom_boxplot(outlier.colour = "black") + labs(title = 'normalized Spectral count vs bait')
p
```

## z-score normalized boxplot count vs condition
boxplot count vs condition
```{r}
p <- ggplot(all_data_dicty, aes(x=condition, y =log2(z_score_spectral)))
p <- p + geom_boxplot(outlier.colour = "black") + labs(title = 'normalized Spectral count vs condition')
p
```

## z-score normalized boxplot boxplot count vs experiment
```{r}
my_data <-all_data_dicty %>% 
  group_by(bait, condition) %>% 
    tidyr::unite(bait_condition, c(bait, condition), sep = "_")

my_data$bait_condition <- as.factor(my_data$bait_condition)

p <- ggplot(my_data, aes(x = bait_condition, y = log2(z_score_spectral)))
p <- p + geom_boxplot(outlier.colour = "black") + labs(title = 'normalized spectral count vs bait + condition')
p + theme(axis.text.x = element_text(angle = 45, hjust = 1))


```


# Venn diagrams 
## Venn diagrams over bait
```{r}
bait <- split(all_data_dicty$uniprot, all_data_dicty$bait)
toy_bait = Venn(bait)
ggvenn(toy_bait, slice = c(1, 2,4))
```

## Venn diagrams over condition
```{r}
condition <- split(all_data_dicty$uniprot, all_data_dicty$condition)
toy_condition = Venn(condition)
ggvenn(toy_condition, slice = c(1, 2,4))
```

## overlap proteins per bait_condition
```{r}
my_data <-all_data_dicty %>% 
  group_by(bait, condition) %>% 
  tidyr::unite(bait_condition, c(bait, condition), sep = "_")


bait_condition <- split(my_data$uniprot, my_data$bait_condition)
toy = Venn(bait_condition)
#setmap(toy) # to many uniprot ids to show on the axis, needs filtering
ggvenn(toy, slice = c(1, 2,4))

```

Method to retrieve the overlapping proteins per experiment
```{r include=FALSE}
# get character vector with overlapping proteins 
overlap(toy, c("Galpha2_starv", "Galpha2_veg", "Galpha8_GDP", "Gbeta1_Normal"))

# use discern to get set difference
# use discern discern(c(a,b), c(c,d)) -> first takes union, then difference
```


# Presence/absence protein
Heatmap for uniprot presence/absence over bait, condition or experiment. Filtered on protein
```{r}
filterd <- filter(all_data_dicty, uniprot == 'sp|P0A9K9|SLYD_ECOLI')

condition <- split(filterd$uniprot, all_data_dicty$condition)
bait <- split(filterd$uniprot, all_data_dicty$bait)

toy_condition_filtered = Venn(condition)
setmap(toy_condition_filtered, element_clustering = FALSE, set_clustering = FALSE)

toy_bait_filtered = Venn(bait)
setmap(toy_bait_filtered, element_clustering = FALSE, set_clustering = FALSE)
```

# Overlapping proteins
```{r}
my_data <-all_data_dicty %>% 
  group_by(bait, condition) %>% 
  tidyr::unite(bait_condition, c(bait, condition), sep = "_")

count(my_data, bait_condition, uniprot) %>%
  spread(bait_condition, n, fill = 0) %>%
  select(-uniprot) %>%
  as.matrix() %>%
  crossprod()


# check value from crossproduct
#                          Galpha2_starv
#RasG1_Normal               279

bait_condition <- split(my_data$uniprot, my_data$bait_condition)
toy = Venn(bait_condition)
length(overlap(toy, c("Galpha2_starv", "RasG1_Normal")))

# same number of proteins reported
```

## Simple heatmap with clustering on overlapping proteins per experiment
```{r}
my_data <- all_data_dicty %>% 
  group_by(bait, condition) %>% 
  tidyr::unite(bait_condition, c(bait, condition), sep = "_")

overlapping <- count(my_data, bait_condition, uniprot) %>%
  spread(bait_condition, n, fill = 0) %>%
  select(-uniprot) %>%
  as.matrix() %>%
  crossprod()

heatmap(overlapping)
```

## ggplot heatmap overlapping proteins per experiment
```{r}
overlap_long <- reshape2::melt(overlapping)
ggplot(overlap_long, aes(x = Var1, y = Var2, fill = value)) +
geom_tile() +
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


# Correlations
## raw data
### bait heatmap on raw spectra correlation
```{r}
pivoted <- all_data_dicty %>% 
  group_by(bait) %>%
  select(uniprot, bait, spectral_count) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = uniprot, values_from = spectral_count, values_fill = 0) %>%
  select(-row) %>%
  summarise_all(funs(sum)) %>%
  tibble::column_to_rownames(var = "bait")

pivoted_t <- t(pivoted)
res_t <- cor(pivoted_t) # experiment vs experiment
heatmap(x = res_t, symm = TRUE)
```

### Condition heatmap on raw spectra correlation
```{r}
pivoted <- all_data_dicty %>% 
  group_by(condition) %>%
  select(uniprot, condition, spectral_count) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = uniprot, values_from = spectral_count, values_fill = 0) %>%
  select(-row) %>%
  summarise_all(funs(sum)) %>%
  tibble::column_to_rownames(var = "condition")

pivoted_t <- t(pivoted)
res_t <- cor(pivoted_t) # experiment vs experiment
heatmap(x = res_t, symm = TRUE)
```

### Experiment heatmap on raw spectra correlation
```{r}
pivoted_experiment <- all_data_dicty %>% 
  group_by(bait, condition) %>%
  select(uniprot, bait, condition, spectral_count) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = uniprot, values_from = spectral_count, values_fill = 0) %>%
  summarise_all(funs(sum)) %>%
  tidyr::unite(bait_condition, c(bait, condition), sep = "_") %>%
  select(-row) %>%
  tibble::column_to_rownames(var = "bait_condition")


pivoted_experiment_t <- t(pivoted_experiment)
res_t_pivoted_experiment <- cor(pivoted_experiment_t)
heatmap(x = res_t_pivoted_experiment, symm = TRUE)
```

## Normalized data

### Bait heatmap on normalized spectra correlation
```{r}
pivoted <- all_data_dicty %>% 
  group_by(bait) %>%
  select(uniprot, bait, z_score_spectral) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = uniprot, values_from = z_score_spectral, values_fill = 0) %>%
  select(-row) %>%
  summarise_all(funs(sum)) %>%
  tibble::column_to_rownames(var = "bait")

pivoted_t <- t(pivoted)
res_t <- cor(pivoted_t) # experiment vs experiment
heatmap(x = res_t, symm = TRUE)
```

### Condition heatmap on normalized spectra correlation
```{r}
pivoted <- all_data_dicty %>% 
  group_by(condition) %>%
  select(uniprot, condition, z_score_spectral) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = uniprot, values_from = z_score_spectral, values_fill = 0) %>%
  select(-row) %>%
  summarise_all(funs(sum)) %>%
  tibble::column_to_rownames(var = "condition")

pivoted_t <- t(pivoted)
res_t <- cor(pivoted_t) # experiment vs experiment
heatmap(x = res_t, symm = TRUE)
```

### Experiment heatmap on normalized spectra correlation
```{r}
pivoted_experiment <- all_data_dicty %>% 
  group_by(bait, condition) %>%
  select(uniprot, bait, condition, z_score_spectral) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = uniprot, values_from = z_score_spectral, values_fill = 0) %>%
  summarise_all(funs(sum)) %>%
  tidyr::unite(bait_condition, c(bait, condition), sep = "_") %>%
  select(-row) %>%
  tibble::column_to_rownames(var = "bait_condition")


pivoted_experiment_t <- t(pivoted_experiment)
res_t_pivoted_experiment <- cor(pivoted_experiment_t)
heatmap(x = res_t_pivoted_experiment, symm = TRUE)
```


# PCA 
## bait on raw spectral_count 
```{r}

pivoted <- all_data_dicty %>% 
  group_by(bait) %>%
  select(uniprot, bait, spectral_count) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = uniprot, values_from = spectral_count, values_fill = 0) %>%
  select(-row) %>%
  summarise_all(funs(sum)) %>%
  tibble::column_to_rownames(var = "bait")
  

wdbc.pr <- prcomp(pivoted)
summary(wdbc.pr)


screeplot(wdbc.pr, type = "l", npcs = 12, main = "Screeplot of the first 12 PCs")
abline(h = 1, col="red", lty=5)
legend("topright", legend=c("Eigenvalue = 1"),
       col=c("red"), lty=5, cex=0.6)
cumpro <- cumsum(wdbc.pr$sdev^2 / sum(wdbc.pr$sdev^2))
plot(cumpro[0:15], xlab = "PC #", ylab = "Amount of explained variance", main = "Cumulative variance plot")

plot(wdbc.pr$x[,1],wdbc.pr$x[,2], xlab="PC1 (50%)", ylab = "PC2 (20%)", main = "PC1 / PC2 - plot")

library("factoextra")
fviz_pca_ind(wdbc.pr, geom.ind = "point", pointshape = 21, 
             pointsize = 2, 
             fill.ind = rownames(pivoted), 
             col.ind = "black", 
             palette = "ucscgb", 
            # addEllipses = TRUE,
             label = "var",
             col.var = "black",
             repel = TRUE,
             legend.title = "bait") +
  ggtitle("2D PCA-plot") +
  theme(plot.title = element_text(hjust = 0.5))
```

## Condition on raw spectral_count 
```{r}
pivoted <- all_data_dicty %>% 
  group_by(condition) %>%
  select(uniprot, condition, spectral_count) %>%
  mutate(row = row_number()) %>%
  tidyr::pivot_wider(names_from = uniprot, values_from = spectral_count, values_fill = 0) %>%
  select(-row) %>%
  summarise_all(funs(sum)) %>%
  tibble::column_to_rownames(var = "condition")
  

wdbc.pr <- prcomp(pivoted)
summary(wdbc.pr)


screeplot(wdbc.pr, type = "l", npcs = 6, main = "Screeplot of the first 6 PCs")
abline(h = 1, col="red", lty=5)
legend("topright", legend=c("Eigenvalue = 1"),
       col=c("red"), lty=5, cex=0.6)
cumpro <- cumsum(wdbc.pr$sdev^2 / sum(wdbc.pr$sdev^2))
plot(cumpro[0:15], xlab = "PC #", ylab = "Amount of explained variance", main = "Cumulative variance plot")

plot(wdbc.pr$x[,1],wdbc.pr$x[,2], xlab="PC1 (87%)", ylab = "PC2 (11%)", main = "PC1 / PC2 - plot")

library("factoextra")
fviz_pca_ind(wdbc.pr, geom.ind = "point", pointshape = 21, 
             pointsize = 2, 
             fill.ind = rownames(pivoted), 
             col.ind = "black", 
             palette = "ucscgb", 
            # addEllipses = TRUE,
             label = "var",
             col.var = "black",
             repel = TRUE,
             legend.title = "condition") +
  ggtitle("2D PCA-plot") +
  theme(plot.title = element_text(hjust = 0.5))
```

## Experiment (bait + condition) on raw spectral_count 
```{r}
my_data <- all_data_dicty %>% 
  group_by(bait, condition) %>% 
  tidyr::unite(bait_condition, c(bait, condition), sep = "_")

pivoted <- my_data %>% 
  select(uniprot, bait_condition, spectral_count) %>%
  group_by(bait_condition) %>%
  pivot_wider(names_from = uniprot, values_from = spectral_count, values_fill = 0) %>%
  summarise_all(funs(sum)) %>%
  tibble::column_to_rownames(var = "bait_condition") 

wdbc.pr <- prcomp(pivoted)
summary(wdbc.pr)


screeplot(wdbc.pr, type = "l", npcs = 15, main = "Screeplot of the first 15 PCs")
abline(h = 1, col="red", lty=5)
legend("topright", legend=c("Eigenvalue = 1"),
       col=c("red"), lty=5, cex=0.6)
cumpro <- cumsum(wdbc.pr$sdev^2 / sum(wdbc.pr$sdev^2))
plot(cumpro[0:15], xlab = "PC #", ylab = "Amount of explained variance", main = "Cumulative variance plot")
#abline(v = 6, col="blue", lty=5)
#abline(h = 0.88759, col="blue", lty=5)
#legend("topleft", legend=c("Cut-off @ PC6"),col=c("blue"), lty=5, cex=0.6)


plot(wdbc.pr$x[,1],wdbc.pr$x[,2], xlab="PC1 (47.4%)", ylab = "PC2 (20.3%)", main = "PC1 / PC2 - plot")

library("factoextra")
fviz_pca_ind(wdbc.pr, geom.ind = "point", pointshape = 21, 
             pointsize = 2, 
             fill.ind = rownames(pivoted), 
             col.ind = "black", 
             palette = "ucscgb", 
            # addEllipses = TRUE,
             label = "var",
             col.var = "black",
             repel = TRUE,
             legend.title = "bait_condition") +
  ggtitle("2D PCA-plot") +
  theme(plot.title = element_text(hjust = 0.5))
```


# Igraph network plot
```{r}
library(igraph)
library(visNetwork)

my_data <- all_data_dicty %>% 
  group_by(bait, condition) %>% 
  tidyr::unite(bait_condition, c(bait, condition), sep = "_")

x <- my_data %>% 
  separate(uniprot, c("a", "b", "c", "protid", "genid", "source"), extra = "merge", fill = "left", sep = '[_|]')
x <- x %>% 
  select(-one_of(c("a", "b", "c")))

# resulting in 17k+ interactions, hairball!, need to filter
new_x <- x %>% 
  filter(spectral_count > 200)

nodes <- as.data.frame(unique(new_x$bait_condition))
nodes$shape = 'triangle'
colnames(nodes) <- c('id', 'shape')

nodes <- nodes %>% 
  add_row(id = unique(new_x$genid), shape = 'circle')

#nodes <- nodes %>% add_row(id = unique(new_x$genid))
nodes$label <- nodes$id

# nodes (shapes) and edges (thickness  = width, use normalized counts) need to be done
# nodes scale (value) on basis of number of interactions
# gene ids, are not complete, <NA> or missing. Check the seperate function!


my_edges <- data.frame(new_x$bait_condition, new_x$genid, new_x$z_score_spectral)
colnames(my_edges) <- c('from', 'to', 'value')

graph <- graph_from_data_frame(my_edges, directed = FALSE)

# color groups based on community detection, method louvain
#cluster <- cluster_louvain(graph)
#cluster_df <- data.frame(as.list(membership(cluster)))
#cluster_df <- as.data.frame(t(cluster_df))
#cluster_df$label <- rownames(cluster_df)
#nodes <- left_join(nodes, cluster_df, by = 'label')
#colnames(nodes)[3] <- 'group'

visNetwork(nodes = nodes, edges = my_edges) %>%
  visOptions(highlightNearest = list(enabled = TRUE, algorithm = "hierarchical", 
   degree = list(from = 1, to = 1)), nodesIdSelection = TRUE) %>%
  visPhysics(stabilization = FALSE, barnesHut = ) %>%
  visLayout(improvedLayout = TRUE)
```

# Demo Viz functions
```{r}
# raw data
plot_spectral_using_bait("Galpha8", 'dicty')
plot_spectral_using_condition("Normal", 'dicty')
plot_spectral_bait_condition_using_uniprot('P36415', 'dicty')
plot_spectral_bait_condition_using_uniprot('GST', 'dicty')
plot_spectral_bait_condition_using_uniprot('Q54YT4', 'dicty')
plot_spectral_bait_condition_using_uniprot('GST', 'neutro')

condition_boxplot_spectral_using_bait("Galpha8", 'dicty')
experiment_boxplot_spectral_using_bait("Galpha8", 'dicty')


# normalized data
#plot_normalized_spectral_using_bait("Galpha8", 'dicty') # we normalized over the total experiments, can we use that to show just bait?
#plot_normalized_spectral_using_condition("Normal", 'dicty') # same here, plots look funny
plot_normalized_spectral_bait_condition_using_uniprot('P36415', 'dicty')

plot_normalized_spectral_bait_condition_using_uniprot('RASC', 'dicty')


#condition_boxplot_normalized_spectral_using_bait("Galpha8", 'dicty') #
experiment_boxplot_normalized_spectral_using_bait("Galpha8", 'dicty')

```

