---
title: "Dicty analysis"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#TODO add root folder using knitr root.dir()
```


```{r message = FALSE, warning=FALSE}
library(ggplot2)

# load the data and visualization functions
source("global.R")
source("../utils/visualization_functions.R")
```

Normalised spectral count per group and total, add columns to dataframe
```{r}
# Z-score normalisation, by group and total
all_data_dicty <- all_data_dicty %>% group_by(bait, condition) %>% mutate(grouped_z_score_spectral = scale(spectral_count))
all_data_dicty <- all_data_dicty %>% mutate(z_score_spectral = scale(spectral_count))
```


Summary statistics Dicty
```{r}
all_data_dicty %>% summarise(Unique_bait = n_distinct(bait))
```


Show the unique baits count
```{r}
all_data_dicty %>% group_by(bait, condition) %>% unite(bait_condition, c(bait, condition), sep = '_') %>% summarise(Unique_bait_condition = n_distinct(bait_condition))
```


Show how many instances (lines) of bait there are
```{r}
table(all_data_dicty$bait)
```

Show how many instances (lines) of condition there are
```{r}
table(all_data_dicty$condition)
```

Show spectral count per bait + condition
```{r}
all_data_dicty %>% group_by(bait, condition) %>% summarise(.groups = 'drop', across(spectral_count, sum))
```

Show normalized spectral count per bait + condition
```{r}
all_data_dicty %>% group_by(bait, condition) %>% summarise(.groups = 'drop', across(z_score_spectral, sum))
# this does not look correct, we shoudld normalize by total spectral count?
```

Show how many uniq uniprot id's are in the set (dicty has 6 chromosomes and 12500 proteins)
```{r}
n_distinct(all_data_dicty$uniprot)
```

uniq proteins per bait_condition
```{r}
bait_condition <- group_by(all_data_dicty, bait, condition)
bait_condition %>% summarise(n_distinct(uniprot))
```


Unique per condition
```{r}
group_by(all_data_dicty, condition) %>% summarise(n_distinct(uniprot))
```

Unique per bait
```{r}
group_by(all_data_dicty, bait) %>% summarise(n_distinct(uniprot))
```


Barplot showing the spectral count for each bait
```{r}
p <- ggplot(all_data_dicty, aes(x=bait, y =spectral_count))
p <- p + geom_col() + labs(title = 'Spectral count vs bait')
p
```


Barplot showing the normalized spectral count for each bait
divided the summed group count by the number of observations in that group so avg per group
```{r}
# get group sizes
group_spectra_normalized <- all_data_dicty %>% group_by(bait) %>% summarise(sum_spectra = sum(spectral_count), group_size = n(), normalized_spectra = sum(spectral_count/n()))


p <- ggplot(group_spectra_normalized, aes(x=bait, y = normalized_spectra))
p <- p + geom_col() + labs(title = 'Normalized Spectral count vs bait')
p
```


Barplot showing the absolute spectral count for each condition
```{r}
p <- ggplot(all_data_dicty, aes(x=condition, y =spectral_count))
p <- p + geom_col() + labs(title = 'Spectral count vs condition')
p
```

Barplot showing the normalized spectral count for each condition:
divided the summed group count by the number of observations in that group
```{r}

# get group sizes
group_spectra_normalized <- all_data_dicty %>% group_by(condition) %>% summarise(sum_spectra = sum(spectral_count), group_size = n(), normalized_spectra = sum(spectral_count/n()))


p <- ggplot(group_spectra_normalized, aes(x=condition, y=normalized_spectra))
p <- p + geom_col() + labs(title = 'Normalized Spectral count vs condition')
p
```



Combined spectral count for bait + condition, sum spectral count per group
```{r}
my_data <-all_data_dicty %>% group_by(bait, condition) %>% 
    # .groups = drop is the default, will get warnings if you not specify explicitly 
    summarise(.groups = 'drop', across(spectral_count, sum)) %>% 
    unite(bait_condition, c(bait, condition), sep = "_")
  
p <- ggplot(my_data, aes(x = bait_condition, y = spectral_count))
p <- p + geom_col() + labs(title = 'spectral count vs bait + condition')
p + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


Combined spectral count for bait + condition, sum normalized spectral count per group
```{r}
# get group sizes
group_spectra_normalized <- all_data_dicty %>% group_by(bait, condition) %>% summarise(sum_spectra = sum(spectral_count), group_size = n(), normalized_spectra = sum(spectral_count/n())) %>% unite(bait_condition, c(bait, condition), sep = "_")


p <- ggplot(group_spectra_normalized, aes(x = bait_condition, y = normalized_spectra))
p <- p + geom_col() + labs(title = 'Normalized spectral count vs bait + condition')
p + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# TODO overlay with uniq number of proteins in the group?
# TODO correct for number of uniq proteins in the group? is every observation unique? (yes, checked)
```{r}
# get group sizes
group_spectra_normalized <- all_data_dicty %>% group_by(bait, condition) %>% summarise(sum_spectra = sum(spectral_count), group_size = n(), normalized_spectra = sum(spectral_count/n()), unique_proteins = n_distinct(uniprot)) %>% unite(bait_condition, c(bait, condition), sep = "_")


p <- ggplot(group_spectra_normalized, aes(x = bait_condition, y = normalized_spectra))
p <- p + geom_col() + labs(title = 'Normalized spectral count vs bait + condition') +
geom_text(aes(label=unique_proteins), vjust=0)

p + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


Boxplot count vs bait
```{r}
p <- ggplot(all_data_dicty, aes(x=bait, y =log10(spectral_count)))
p <- p + geom_boxplot(outlier.colour = "black") + labs(title = 'Spectral count vs bait')
p
```


Boxplot count vs condition
```{r}
p <- ggplot(all_data_dicty, aes(x=condition, y =log10(spectral_count)))
p <- p + geom_boxplot(outlier.colour = "black") + labs(title = 'Spectral count vs condition')
p
```


Boxplot combined spectral count for bait + condition
```{r}
my_data <-all_data_dicty %>% group_by(bait, condition) %>% 
    unite(bait_condition, c(bait, condition), sep = "_")

my_data$bait_condition <- as.factor(my_data$bait_condition)
p <- ggplot(my_data, aes(x = bait_condition, y = log10(spectral_count)))

p <- p + geom_boxplot(outlier.colour = "black") + labs(title = 'spectral count vs bait + condition')

p + theme(axis.text.x = element_text(angle = 45, hjust = 1))



```




# TODO, check why we have the same numbers for some conditions or baits
#filter(all_data_dicty, bait == "Gbeta2")
#filter(all_data_dicty, bait == "Gbeta1")
#Shows the same number of rows retrieved, counts look different for now....


boxplot count vs condition
```{r}
p <- ggplot(all_data_dicty, aes(x=condition, y =log2(z_score_spectral)))
p <- p + geom_boxplot(outlier.colour = "black") + labs(title = 'normalized Spectral count vs bait')
p
```


boxplot count vs condition
```{r}
p <- ggplot(all_data_dicty, aes(x=condition, y =log2(z_score_spectral)))
p <- p + geom_boxplot(outlier.colour = "black") + labs(title = 'normalized Spectral count vs condition')
p
```


boxplot count + condition
```{r}
my_data <-all_data_dicty %>% group_by(bait, condition) %>% 
    unite(bait_condition, c(bait, condition), sep = "_")

my_data$bait_condition <- as.factor(my_data$bait_condition)
p <- ggplot(my_data, aes(x = bait_condition, y = log2(z_score_spectral)))

p <- p + geom_boxplot(outlier.colour = "black") + labs(title = 'normalized spectral count vs bait + condition')

p + theme(axis.text.x = element_text(angle = 45, hjust = 1))


```






Summary statistics Neutro
```{r}
table(all_data_neutro$bait)
table(all_data_neutro$condition)


```


```{r}

plot_spectral_bait_condition_using_uniprot('P36415', 'dicty')
plot_spectral_bait_condition_using_uniprot('GST', 'dicty')
plot_spectral_bait_condition_using_uniprot('GST', 'neutro')
plot_spectral_bait_condition_using_uniprot('Q54YT4', 'dicty')

```
